End-User Software Customization by Direct Manipulation of Tabular Data

### 2020-08-09 initial-submission.tex
record initial submission files

\begin{itemize}
\tightlist
\item
  \emph{Customization by direct manipulation}: We discuss further how
  table-driven customization relates to a gentle slope of customization.
  We suggest that an important point on this slope is the ability to
  customize an application by directly seeing and changing its data,
  rather than by writing imperative scripts.
\item
  \emph{Semantic wrappers}: Typically, tools that don't rely on official
  extension APIs resort to offering low-level APIs for customization.
  Instead, we propose a community-maintained library of semantic
  wrappers around existing applications, enabling end users to work with
  domain objects rather than low-level representations.
\end{itemize}

@@

One aspect of directness that we have chosen not to pursue in Wildcard
is enabling customization in closer proximity to the original user
interface elements, as explored by tools like Scotty
\citep{eagan2011}---all customization occurs in a separate panel next to
the original UI. While closer proximity might be helpful, we have found
that augmenting the original UI with a distinct, additional
representation provides a more consistent experience across all
applications, and clearly shows what structured data is available to
work with.

@@

This paper extends work reported in a workshop paper by Litt and Jackson
\citep{litt2020} which presented an earlier version of the Wildcard
extension. We have substantially extended their work in this paper by
creating the table adapter abstarction, reimplementing the internals of
Wildcard around that abstraction, evaluating the system on many more
websites, and by characterizing the design of the system in much more
detail than in their workshop paper.

@@

More recently, web customization tools have aimed to enable end users to
modify web interfaces without programming.

Sifter \citep{huynh2006} enables end users to sort and filter lists of
data obtained by web scraping, much like Wildcard's sorting features.
The main difference between the systems is that table-driven
customization has many other use cases besides sorting and filtering.
Also, Sifter involves end users in a semi-automated data extraction
process, rather than having programmers create wrappers. This provides
coverage of more websites, but at the expense of complicating the end
user experience. We could integrate end user scraping techniques in
Wildcard in the future, but we believe that when possible it's valuable
for end users to have a customization experience decoupled from the
challenge of web scraping the underlying data. Sifter also implements
scraping across multiple pages, a valuable feature for sorting and
filtering that isn't present in Wildcard.

@@

\hypertarget{direct-manipulation-programming-interfaces.}{%
\subsection{Direct manipulation programming
interfaces.}\label{direct-manipulation-programming-interfaces.}}

Another relevant area involves spreadsheets and visual query systems. We
take inspiration from these tools in our work, but apply them in a
different domain: customizing existing software applications, rather
than interacting with databases or constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data. SIEUFERD by Bakke and Karger
\citep{bakke2016} is one such recent system, and their paper presents a
survey of many other similar tools. Our work is particularly influenced
by the authors' observation that direct manipulation requires that the
user manipulate the results of a database query rather than the query
itself, and that the user must see intermediate results at every step of
constructing the query. SIEUFERD's interface supports a far wider range
of queries than Wildcard, but the core principles of the user interface
are similar. \href{https://airtable.com/}{Airtable} is another example
of a modern commercial product that offers spreadsheet-like interaction
with a relational database.

>>>>

\begin{itemize}
\tightlist
\item
  \emph{Customization by direct manipulation}: We discuss further how
  table-driven customization relates to a gentle slope of customization.
  We suggest that an important point on this slope is the ability to
  customize an application by directly seeing and changing its data,
  rather than by writing imperative scripts.
\item
  \emph{Semantic wrappers}: Typically, tools that don't rely on official
  extension APIs resort to offering low-level APIs for customization.
  Instead, we propose a community-maintained library of semantic
  wrappers around existing applications, enabling end users to work with
  domain data rather than low-level representations.
\end{itemize}

@@

One aspect of directness that we have chosen not to pursue in Wildcard
is enabling customization in closer proximity to the original user
interface elements, as explored by tools like Scotty
\citep{eagan2011}---all customization occurs in a separate panel next to
the original UI. While closer proximity might be helpful, we have found
that augmenting the original UI with a distinct, additional
representation provides a more consistent experience across all
applications, and clearly shows what structured data is available to
work with. We also emphasize the mapping between the representations by
highlighting content in the original page, similiar to the way that
browser developer tools highlight the currently selected element in the
DOM inspector in the original page.

@@

This paper extends work reported in a workshop paper by Litt and Jackson
\citep{litt2020} which presented an early version of the Wildcard
extension. We have substantially extended their work in this paper by
creating the table adapter abstraction, reimplementing the internals of
Wildcard around that abstraction, evaluating the system on many more
websites and use cases, and by characterizing the design of the system
in much more detail than in their workshop paper.

@@

More recently, web customization tools have aimed to enable end users to
modify web interfaces without programming. Sifter \citep{huynh2006}
enables end users to sort and filter lists of data obtained by web
scraping, much like Wildcard's sorting features. The main difference
between the systems is that table-driven customization has many other
use cases besides sorting and filtering. Also, Sifter involves end users
in a semi-automated data extraction process, rather than having
programmers create wrappers. This provides coverage of more websites,
but at the expense of complicating the end user experience. We could
integrate end user scraping techniques in Wildcard in the future, but we
believe that when possible it's valuable for end users to have a
customization experience decoupled from the challenge of web scraping
the underlying data. Sifter also implements scraping across multiple
pages, a valuable feature for sorting and filtering that isn't present
in Wildcard.

@@

\hypertarget{spreadsheets-and-visual-query-interfaces}{%
\subsection{Spreadsheets and visual query
interfaces}\label{spreadsheets-and-visual-query-interfaces}}

Another relevant area involves spreadsheets and visual query interfaces.
We take inspiration from these tools in our work, but apply them in a
different domain: customizing existing software applications, rather
than interacting with databases or constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data. SIEUFERD by Bakke and Karger
\citep{bakke2016} is one such recent system, and their paper presents a
survey of many other similar tools. Our work is particularly influenced
by the authors' observation that a user should be able to modify queries
by interacting with the results of the query rather than some
representation of the query itself. SIEUFERD's interface supports a far
more general range of queries than Wildcard, but the core principles of
the user interface are the same. \href{https://airtable.com/}{Airtable}
is another example of a modern commercial product that offers
spreadsheet-like interaction with a relational database.
